---
title: "trackSurvey: Survey Progress Visualizer"
author: "Kevin L. Stierhoff"
format: html
editor: visual
---

```{r setup,echo=F,message=F,warning=F,error=F,include=F}
# Install and load pacman (library management package)
if (!require("pacman")) install.packages("pacman")
if (!require("pak")) install.packages("pak")

# Install and load required packages from CRAN ---------------------------------
pacman::p_load(tidyverse,raster,cowplot,DBI,odbc,RSQLite,scatterpie,ggspatial,vroom,
               readxl,lubridate,rnaturalearth,sf,here,naniar,mapview,xts,terra,
               shadowtext,knitr,ggrepel,fs,plotly,leaflet,leaflet.extras,bookdown,
               htmltools,leafem,DT,rworldmap,rworldxtra,lwgeom,htmlwidgets,httr,
               ncdf4,rerddap,patchwork,kableExtra,xml2)

# Install and load required packages from Github -------------------------------
if (!require("atm")) pak::pkg_install("SWFSC/atm")
if (!require("surveyR")) pak::pkg_install("SWFSC/surveyR")
pacman::p_load_gh("SWFSC/atm")
pacman::p_load_gh("SWFSC/surveyR")

# rnaturalearth data
pacman::p_load_gh("ropenscilabs/rnaturalearthdata")
pacman::p_load_gh("ropenscilabs/rnaturalearthhires")

# Define method of table generation (whether kable or xtable) for best formatting
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
doc.name <- knitr::current_input()
if (is.null(doc.type)) {doc.type <- "html"}
if (is.null(doc.name)) {doc.name <- "plotSurvey.Rmd"}

# global knitr chunk options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, 
  fig.align = "center", dev = "png", dev.args = list(type = "cairo"), dpi = 150
)
knitr::opts_chunk$set(cache = FALSE)

# Determine global knitr table format
if (doc.type == "latex") {
  knitr.format <- "latex"
} else {
  knitr.format <- "html" 
}
```

```{r user-input, include=FALSE}
# Get project name from directory
prj.name <- last(unlist(str_split(here(), "/")))

# Get all settings files
settings.files <- dir(here("Doc/settings"))

# Source survey settings file
prj.settings <- settings.files[str_detect(settings.files, paste0("settings_", prj.name, ".R"))]

source(here("Doc/settings", prj.settings))

# Set allowances for transit days during the survey
das.survey.transit <- 1
```

```{r processing-controls}
# Control script behavior (usually T)
get.nav            <- T # Download NOAA Ship nav data
save.figs          <- T # Save figures
copy.files         <- T # Copy data files from data to estimATM directory
overwrite.files    <- T # Overwrite existing files when copying (not CSV, see below)

# Control script behavior (often F)
resize.map         <- F # Resize map during survey; if T, uses anticipated bounds of survey area
```

```{r config-project}
dir_create(here(), c("Output", "Figs"))
```

```{r copy-files, include=F}
if (copy.files) {
  # Create data directories
  dir_create(here("Data", c("SCS")))
  
  if (nav.source == "SCS") {
    # Copy SCS files
    source(here("Code/copy_files_scs.R"))
  }
}
```

```{r process-nav,include=F}
# Source code to get nav data from ERDDAP, RosePoint, or SCS
if (nav.source == "ERDDAP") {
  source(here("Code/get_nav_erddap.R"))
} else if (nav.source == "SCS") {
  source(here("Code", scs.nav.script))
} else if (nav.source == "RosePoint") {
  source(here("Code/get_nav_rosepoint.R"))
}

# if (get.nav.sh) {
#   source(here("Code/get_nav_erddap_SH.R"))
# }

# Remove non-survey legs
nav.paths.sf <- nav.paths.sf %>% 
  filter(!leg == "Leg NA", !is.na(Leg))

# Read transect waypoints
wpts <- read_csv(here("Data/Nav", wpt.filename))

# Convert planned transects to sf; CRS = crs.geog
wpts.sf <- wpts %>% 
  filter(Type %in% wpt.types, Region %in% wpt.regions) %>% 
  st_as_sf(coords = c("Longitude","Latitude"), crs = crs.geog) %>% 
  mutate(
    label = paste("Transect", Transect),
    popup = paste('<b>Transect:</b>', Transect, Type)
  )

transects.sf <- wpts.sf %>% 
  group_by(Type, Transect, Region) %>% 
  summarise(do_union = FALSE) %>% 
  st_cast("LINESTRING") %>% 
  ungroup() %>% 
  mutate(
    distance = round(as.numeric(st_length(.))/1852,1),
    label    = paste("Transect", Transect),
    popup    = paste('<b>Transect:</b>', Transect, Type, '<br/>',
                     'Distance:', distance, 'nmi<br/>')) 

if (exists("tx.breaks")) {
  transects.sf <- transects.sf %>% 
    mutate(Leg = as.factor(cut(Transect, tx.breaks, labels = FALSE)))
  
  wpts.sf <- wpts.sf %>% 
    mutate(Leg = as.factor(cut(Transect, tx.breaks, labels = FALSE)))
}

# Get most recent vessel position for plotting
nav.now <- tail(nav.sf, 1) %>% 
  mutate(time_local = with_tz(time, tzone = "America/Los_Angeles")) %>% 
  mutate(label = paste("Last position:", time, "UTC"),
         popup = paste0('<b>Vessel name: </b>', survey.vessel, '<br/>',
                        '<b>Last position (UTC): </b>', time, ' UTC<br/>',
                        '<b>Last position (Local): </b>', time_local, ' PT<br/>')) %>% 
  # Calculate legs using the actual leg dates
  mutate(leg = cut(as.numeric(date(time)), 
                   as.numeric(leg.ends),
                   labels = FALSE)) %>% 
  filter(!is.na(leg), leg %% 2 != 0) 

# Set padding around data  
if (resize.map) {
  # Use nav data to resize map to survey progress
  map.bounds <- nav.paths.sf %>% 
    st_transform(crs = crs.proj) %>%
    st_bbox()  
} else {
  # Use transects to set map to survey extent
  map.bounds <- transects.sf %>%
    st_transform(crs = crs.proj) %>%
    st_bbox()  
}
```

```{r create-map-labels}
# Create acoustic transect labels for maps
tx.labels.tmp <- wpts %>% 
  filter(Type %in% c("Compulsory","Adaptive"),
         Region %in% wpt.regions) %>% 
  group_by(Type, Transect) %>% 
  summarise(
    transect.name = Transect[1],
    start.lat = Latitude[which.max(Longitude)],
    start.long = max(Longitude),
    end.lat = Latitude[which.min(Longitude)],
    end.long = min(Longitude),
    brg = 90 - swfscMisc::bearing(end.lat,end.long,start.lat,start.long)[1]) %>% 
  ungroup()

tx.end.labels <- tx.labels.tmp %>% 
  filter(start.lat < 48.54116) %>% 
  select(Type, transect.name, lat = end.lat, long = end.long, brg) 

tx.start.labels <- tx.labels.tmp %>% 
  filter(start.lat >= 48.54116) %>% 
  select(Type, transect.name, lat = start.lat, long = start.long, brg) %>% 
  rbind(tx.end.labels)

tx.labels <- project_df(tx.start.labels, to = crs.proj)
```

```{r create-basemap,include=F}
# Configure base map options -----------------
# Import landmarks
locations <- filter(read.csv(here("Data/Map/locations.csv")), name %in% label.list) %>%
  project_df(to = crs.proj)

# Get state data
states <- ne_states(country = 'United States of America', returnclass = 'sf')
ca     <- filter(states, name == "California")

# Get countries
countries <- ne_countries(scale = "large", returnclass = "sf") %>%
  filter(subregion %in% c("Northern America","Central America"))

# Read bathy contours shapefile 
bathy <- st_read(here("Data/GIS/bathy_contours.shp")) %>% 
  st_transform(crs.geog) %>% 
  rename(Depth = Contour)

# Determine map aspect ratio and set height and width
map.aspect <- (map.bounds$xmax - map.bounds$xmin)/(map.bounds$ymax - map.bounds$ymin)
map.height <- 10
map.width  <- map.height*map.aspect

# Create base map
base.map <- get_basemap(nav.paths.sf, states, countries, locations, bathy, map.bounds, crs = crs.proj) 

# Save the basemap
ggsave(base.map, file = here("Figs/fig_basemap.png"), 
       height = map.height, width = map.width)

save(base.map, file = here("Data/Map/basemap.Rdata"))
```

```{r get-spatial-files,include=F}
# Import shapefiles ------------------------------------------------
# Read CA State Waters shapefile
ca_waters <- st_read(here("Data/GIS/ca_state_waters.shp")) %>% 
  st_transform(crs.geog)

or_waters <- st_read(here("Data/GIS/or_state_waters.shp")) %>% 
  st_transform(crs.geog)

# Read CA MPAs shapefile
ca_mpas <- st_read(here("Data/GIS/ca_mpas.shp")) %>% 
  st_transform(crs.geog) %>% 
  mutate(MPA = paste(NAME, Type))

or_mpas <- st_read(here("Data/GIS/or_mpas.shp")) %>% 
  st_transform(crs.geog) 

wc_mpas <- st_read(here("Data/GIS/wc_mpas.shp")) %>% 
  st_transform(crs.geog) 

nms_bounds <- st_read(here("Data/GIS/national_marine_sanctuaries.shp"))

# Read MPA shapefiles  -----------------------------------------
# Canada
scott_islands_mpa <- st_read(here("Data/GIS/scott_islands_mpa_polygon.shp")) %>% 
  st_transform(crs.geog) %>% 
  mutate(MPA = "Scott Islands Protected Marine Area") %>% 
  st_difference(filter(countries, name == "Canada"))

pacific_rim_mpa <- st_read(here("Data/GIS/pacific_rim_reserve_polygon.shp")) %>% 
  st_transform(crs.geog) %>% 
  mutate(MPA = "Pacific Rim National Park Reserve") 

# Mexico
baja_mpas <- st_read(here("Data/GIS/baja_mpas.shp")) %>% 
  st_transform(crs.geog)

mex_mpas <- st_read(here("Data/GIS/mex_mpas.shp")) %>% 
  st_transform(crs.geog)

# Read CA MPAs shapefile
eez_usa <- st_read(here("Data/GIS/eez_us.shp")) %>% 
  st_transform(crs.geog)
eez_can <- st_read(here("Data/GIS/eez_canada.shp")) %>% 
  st_transform(crs.geog)
eez_mex <- st_read(here("Data/GIS/eez_mex.shp")) %>% 
  st_transform(crs.geog)
```

```{r plot-survey-progress}
# Load backscatter data
load(here("Output/nasc_cps.Rdata"))

# Add leg information to nasc data
nasc.cps <- nasc.cps %>% 
  # Calculate legs using the actual leg dates
  mutate(leg = cut(as.numeric(date(datetime)), 
                   leg.breaks,
                   labels = FALSE)) %>% 
  filter(!is.na(leg))
  # filter(!is.na(leg), leg %% 2 != 0)

# Summarize latitude by vessel and date
nasc.summ.all <- nasc.cps %>% 
  mutate(date.pdt = with_tz(datetime, tzone = "America/Los_Angeles")) %>%
  mutate(date = date(date.pdt),
         vessel.name = survey.vessel.primary) %>% 
  mutate(vessel = case_when(
    vessel.name == "SH"  ~ "Shimada",
    vessel.name == "RL"  ~ "Lasker",
    vessel.name == "LM"  ~ "Lisa Marie",
    vessel.name == "LBC" ~ "Long Beach Carnage")) %>% 
  group_by(vessel, date) %>% 
  summarise(lat = mean(lat)) %>% 
  mutate(key = paste(vessel, date))

# Summarize nautical miles sampled by vessel and date
nasc.summ.all.dist <- nasc.cps %>% 
  mutate(date.pdt = with_tz(datetime, tzone = "America/Los_Angeles")) %>%
  mutate(date = date(date.pdt),
         vessel.name = survey.vessel.primary) %>% 
  mutate(vessel = case_when(
    vessel.name == "SH"  ~ "Shimada",
    vessel.name == "RL"  ~ "Lasker",
    vessel.name == "LM"  ~ "Lisa Marie",
    vessel.name == "LBC" ~ "Long Beach Carnage")) %>% 
  group_by(vessel, date) %>% 
  summarise(dist = n() * nasc.interval / 1000) %>% 
  mutate(key = paste(vessel, date),
         dist.cum = cumsum(dist))

# Summarize underway nav data by vessel and date
nav.summ.fsv <- nav %>% 
  mutate(time.pdt = with_tz(time, tzone = "America/Los_Angeles")) %>% 
  mutate(vessel = survey.vessel,
         date = date(time.pdt)) %>% 
  group_by(vessel, date) %>% 
  summarise(lat = mean(lat),
            SOG = round(mean(SOG, na.rm = TRUE),1)) %>% 
  mutate(key = paste(vessel, date)) %>% 
  mutate(sampling = case_when(
    key %in% unique(nasc.summ.all$key) ~ TRUE,
    TRUE ~ FALSE))
 
# Combine all nav data
nav.summ.all <- nav.summ.fsv #%>% bind_rows(nav.summ.lbc) %>% bind_rows(nav.summ.lm) 

# Format leg breaks
leg.breaks <- data.frame(
  # Get leg end dates
  date = ymd(leg.ends[c(1, seq(2, length(leg.ends), 2))]),
  # Get leg labels
  leg = c("Start", paste("Leg", seq(1, (length(leg.ends)/2), 1)))
)[1:n_distinct(transects.sf$Leg)+1, ]

# Create landmarks data.frame
landmarks <- data.frame(
  lat = c(36.646, 40.50),
  name = c("Monterey","Cape Mendocino"),
  date = mean(c(min(leg.ends),max(leg.ends))))

# Anticipated progress through the transect plan
tx.breaks <- c(min(transects.sf$Transect[transects.sf$Type == "Compulsory"]),
               tx.breaks[2:length(tx.breaks)-1],
               max(transects.sf$Transect[transects.sf$Type == "Compulsory"]))

wpt.summ <- wpts %>% 
  filter(Type == "Compulsory", Region %in% wpt.regions) %>%
  group_by(Transect) %>% 
  slice(which.max(Longitude))

if (file.exists(here("Data/Nav", paste0("transect_plan_", survey.name, ".csv")))) {
  tx.plan <- read_csv(here("Data/Nav", paste0("transect_plan_", survey.name, ".csv"))) %>% 
    mutate(date = mdy(date),
           Transect = floor(tx.plan)) %>% 
    inner_join(select(wpt.summ, Transect, Latitude))
}

# Compute cumulative transect distances
tx.summ <- transects.sf %>% 
  filter(Type %in% c("Compulsory","Adaptive")) %>% 
  mutate(dist = distance*1.852,
         dist.cum = cumsum(dist))

# Add cumulative distances to transect plan
tx.plan <- tx.plan %>% 
  inner_join(select(tx.summ, Transect, dist, dist.cum)) 

# Calculate goal transects
tx.goals <- wpts %>% 
  filter(Transect %in% tx.breaks, Type %in% c("Compulsory")) %>% 
  group_by(Transect) %>% 
  slice(which.max(Longitude)) %>% 
  ungroup() %>% 
  mutate(label = paste("Leg", seq_along(Transect)-1),
         date = min(leg.ends) - days(10)) %>% 
  mutate(label = case_when(
    label == "Leg 0" ~ "Start",
    TRUE ~ label)) %>% 
  # Add cumulative distances to transect goals
  left_join(select(tx.summ, Transect, dist, dist.cum))

# Plot survey progress
vessel.progress.plot <- ggplot() +  #nav.summ.all, aes(date, lat, group = vessel, colour = vessel)
  geom_line(linewidth = 1, linetype = "dashed") + 
  # planned transects
  geom_line(data = tx.plan, aes(date, Latitude), inherit.aes = FALSE,
            linewidth = 1, linetype = "dashed", colour = "gray50") +
  geom_point(data = tx.plan, aes(date, Latitude), inherit.aes = FALSE,
             size = 2, shape = 21, colour = "gray50", fill = "white") +
  # # planned transects - revised
  # geom_line(data = tx.plan2, aes(date, Latitude), inherit.aes = FALSE,
  #           linewidth = 1, linetype = "dashed", colour = "blue") +
  # geom_point(data = tx.plan2, aes(date, Latitude), inherit.aes = FALSE,
  #            size = 2, shape = 21, colour = "blue", fill = "white") +
  geom_vline(xintercept = leg.breaks$date, linetype = "dashed") +
  geom_hline(yintercept = tx.goals$Latitude, linetype = "dashed") +
  geom_text(data = leg.breaks, aes(date, 31, label = leg), inherit.aes = FALSE) +
  geom_text(data = tx.goals, aes(date, Latitude + 0.4, label = label), inherit.aes = FALSE) +
  # geom_text(data = landmarks, aes(date, lat+0.25, label = name), inherit.aes = FALSE) +
  geom_line(data = nav.summ.all, aes(date, lat, group = vessel, colour = vessel)) + 
  # geom_line(data = nav.summ.all, aes(date, lat, group = vessel, colour = vessel),
  #           linewidth = 1, linetype = "dashed") + 
  geom_point(data = nav.summ.all, 
             aes(date, lat, group = vessel, colour = vessel, fill = sampling), 
             size = 2, shape = 21) +
  geom_point(data = filter(nav.summ.all, SOG < 1), 
             aes(date, lat, group = vessel, colour = vessel), 
             fill = "blue", size = 2, shape = 21, show.legend = FALSE) +
  scale_colour_discrete(name = "Vessel") +
  scale_fill_manual(name = "Sampling", values = c("TRUE" = 'black', "FALSE" = 'white')) +
  scale_x_date(date_breaks = "10 days") + 
  labs(x = "Date", y = "Latitude") + 
  theme_bw()

ggsave(vessel.progress.plot, 
       filename = here("Figs/fig_vessel_progress.png"), 
       height = 7, width = 12)
```

```{r plot-survey-progress-distance}
# Plot survey progress
vessel.progress.plot.dist <- ggplot() + 
  geom_line(linewidth = 1, linetype = "dashed") + 
  # planned transects
  geom_line(data = tx.plan, aes(date, dist.cum), inherit.aes = FALSE,
            linewidth = 1, linetype = "dashed", colour = "gray50") +
  geom_point(data = tx.plan, aes(date, dist.cum), inherit.aes = FALSE,
             size = 2, shape = 21, colour = "gray50", fill = "white") +
  geom_vline(xintercept = leg.breaks$date, linetype = "dashed") +
  geom_hline(yintercept = tx.goals$dist.cum, linetype = "dashed") +
  geom_text(data = leg.breaks, aes(date, 500, label = leg), inherit.aes = FALSE) +
  geom_text(data = tx.goals, aes(date, dist.cum + 400, label = label), inherit.aes = FALSE) +
  geom_line(data = nasc.summ.all.dist, aes(date, dist.cum, group = vessel, colour = vessel)) + 
  geom_point(data = nasc.summ.all.dist, 
             aes(date, dist.cum, group = vessel, colour = vessel), 
             size = 2, shape = 21, fill = "white") +
  scale_colour_discrete(name = "Vessel") +
  scale_x_date(date_breaks = "10 days") + 
  labs(x = "Date", y = "Cumulative distance (km)") + 
  theme_bw()

ggsave(vessel.progress.plot.dist, 
       filename = here("Figs/fig_vessel_progress_distance.png"), 
       height = 7, width = 12)
```

```{r create-static-maps}
nasc.cps.sf <- nasc.cps %>%
  st_as_sf(coords = c("long","lat"), crs = crs.geog) %>% 
  group_by(transect, leg) %>% 
  summarise(do_union = FALSE) %>% 
  st_cast("LINESTRING") %>% 
  ungroup()

# Define manual color palette
legPalette <- c("#FF0000", "#FFA500", "#008000", 
                "#0000FF", "#8A2BE2", "#FFFF00")[1:n_distinct(transects.sf$Leg)]

# Map planned transects
map.plan <- base.map + 
  geom_sf(data = filter(transects.sf, !is.na(Leg), Type %in% c("Compulsory","Adaptive")), 
          aes(colour = factor(Leg)),
          linewidth = 1) + 
  # Plot acoustic transect labels N of Cape Flattery
  geom_shadowtext(data = tx.labels,
                  aes(X, Y, label = transect.name,
                      angle = brg),
                  size = 2.5,
                  fontface = 'bold.italic',
                  colour = "black",
                  bg.colour = "white") +
  scale_colour_manual(name = "Leg", values = legPalette) +
  labs(title = "Projected progress",
       subtitle = "Based on start-of-survey estimates") +
  coord_sf(crs = crs.proj, # CA Albers Equal Area Projection
           xlim = c(map.bounds["xmin"], map.bounds["xmax"]), 
           ylim = c(map.bounds["ymin"], map.bounds["ymax"]))

# Map progress
map.progress <- base.map + 
  geom_sf(data = filter(transects.sf, !is.na(Leg), Type %in% c("Compulsory","Adaptive")), 
          colour = "gray50",
          linewidth = 0.5, linetype = "dashed") + 
  geom_sf(data = nav.paths.sf, aes(colour = factor(Leg)),
          linewidth = 0.5, alpha = 0.4) + 
  geom_sf(data = nasc.cps.sf, aes(colour = factor(leg)),
          linewidth = 1) + 
  geom_sf(data = nav.now, aes(colour = factor(Leg)), 
          shape = 23, size = 3, fill = "white") + 
  # Plot acoustic transect labels N of Cape Flattery
  geom_shadowtext(data = tx.labels,
                  aes(X, Y, label = transect.name,
                      angle = brg),
                  size = 2.5, 
                  fontface = 'bold.italic',
                  colour = "black",
                  bg.colour = "white") +
  scale_colour_manual(name = "Leg", values = legPalette) +
  labs(title = "Actual progress",
       subtitle = "Based on data collected") +
  coord_sf(crs = crs.proj, # CA Albers Equal Area Projection
           xlim = c(map.bounds["xmin"], map.bounds["xmax"]), 
           ylim = c(map.bounds["ymin"], map.bounds["ymax"]))

# Arrange figures
map.plan.comp <- map.plan + map.progress

ggsave(map.plan.comp, filename = here("Figs/fig_compare_survey_progress.png"),
       height = 10, width = 10)
```

```{r create-leaflet-map, out.width="100%", out.height="8in"}
# Leaflet options
# https://rstudio.github.io/leaflet/

# Info on tile caching
# https://bhaskarvk.github.io/leaflet.extras/reference/TileCaching.html

# Set padding around data  
imap.bounds <- map_bounds(nav$lat, nav$long, 0.1) 

# Create color palette for MPAs
## CA and OR MPAs
all.mpa.types <- as.factor(c(unique(as.character(ca_mpas$Type)), 
                             unique(as.character(or_mpas$Label))))

mpaPal  <- colorFactor(topo.colors(10), all.mpa.types)

## All West Coast MPAs
mpaPalWC <- colorFactor(topo.colors(10), unique(wc_mpas$Design))

## National Marine Sanctuaries
mpaPalNMS  <- colorFactor(topo.colors(10), as.factor(unique(nms_bounds$NAME)))

# Create color palette for Mexico MPAs
all.mpa.types.mex <- as.factor(c(unique(as.character(baja_mpas$NOMBRE)),
                                 unique(as.character(mex_mpas$NOMBRE))))

mpaPalMex  <- colorFactor(topo.colors(10), all.mpa.types.mex)

# Create color palette for planned transects
txPal    <- colorFactor(wpt.colors, wpt.types)
# Define color palette for legs
txPalLeg <- colorFactor(legPalette,
                        sort(unique(transects.sf$Leg)))

# Create or load oceanIcons
if(!file.exists(here("Images/oceanIcons.Rdata"))) {
  source(here("Code/create_ocean_icons.R"))
} else {
  load(here("Images/oceanIcons.Rdata"))  
}

# Plotting preferences -------------------------------------------------------
# Leaflet tile options; set both to T if caching

# Create leaflet map
i.map <- leaflet() %>% 
  # Enable tile caching
  enableTileCaching() %>%
  # Add provider tiles; # http://leaflet-extras.github.io/leaflet-providers/preview/index.html
  addProviderTiles(providers$Esri.NatGeoWorldMap, group = "Esri.NatGeoWorldMap",
                   options = tileOptions(useCache = useCachedTile,
                                         crossOrigin = useCrossOrigin)) %>%
  addProviderTiles(providers$Esri.OceanBasemap, group = "Esri.OceanBasemap",
                   options = tileOptions(useCache = useCachedTile,
                                         crossOrigin = useCrossOrigin)) %>%
  addProviderTiles(providers$Esri.WorldStreetMap, group = "Esri.WorldStreetMap",
                   options = tileOptions(useCache = useCachedTile,
                                         crossOrigin = useCrossOrigin)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Esri.WorldImagery",
                   options = tileOptions(useCache = useCachedTile,
                                         crossOrigin = useCrossOrigin)) %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik, group = "OpenStreetMap.Mapnik",
                   options = tileOptions(useCache = useCachedTile,
                                         crossOrigin = useCrossOrigin)) %>%
  # Add EEZs
  addPolylines(data = eez_usa, color = "#000414", weight = 3, 
               label = "EEZ-U.S.", group = "Exclusive Economic Zones") %>% 
  addPolylines(data = eez_can, color = "#000414", weight = 3, 
               label = "EEZ-Canada", group = "Exclusive Economic Zones") %>% 
  addPolylines(data = eez_mex, color = "#000414", weight = 3, 
               label = "EEZ-Mexico", group = "Exclusive Economic Zones") %>% 
  # Add bathymetry contours
  addPolylines(data = bathy, color = "white", weight = 2, 
               label = ~paste(Depth, "m"), group = "Bathymetry Contours") %>% 
  # Add National Marine Sanctuaries
  addPolygons(data = nms_bounds, color = "#000414", weight = 2, fillColor = ~mpaPal(NAME),
              fillOpacity = 0.1, label = ~htmlEscape(NAME), group = "U.S. MPAs") %>%
  # Add State waters
  addPolygons(data = or_waters, weight = 2, fillColor = "transparent", 
              opacity = 0.75,
              label = ~htmlEscape("OR State Waters"),
              group = "State Waters") %>%
  addPolygons(data = ca_waters, weight = 2, fillColor = "transparent",
              opacity = 0.75,
              label = ~htmlEscape("CA State Waters"),
              group = "State Waters") %>%
  # Add West Coast MPAs
  addPolygons(data = wc_mpas, color = "#000414", weight = 2, fillColor =  ~mpaPalWC(Design),
              fillOpacity = 0.25, label = ~htmlEscape(Site_Nm), group = "State MPAs") %>%
  # Add Canadian MPAs
  addPolygons(data = scott_islands_mpa, color = "#000414", weight = 2, 
              label = ~htmlEscape(MPA), group = "Canada MPAs") %>%
  addPolygons(data = pacific_rim_mpa, color = "#000414", weight = 2, 
              label = ~htmlEscape(MPA), group = "Canada MPAs") %>%
  # Add Mexico MPAs
  # Baja
  addPolygons(data = baja_mpas, color = "#000414", weight = 2, fillColor =  ~mpaPalMex(NOMBRE),
              fillOpacity = 0.25, label = ~htmlEscape(NOMBRE), group = "Mexico MPAs") %>%
  # All of Mexico
  addPolygons(data = mex_mpas, color = "#000414", weight = 2, fillColor =  ~mpaPalMex(NOMBRE),
              fillOpacity = 0.25, label = ~htmlEscape(NOMBRE), group = "Mexico MPAs") %>%
  # Add core planned transects
  addPolylines(data = filter(transects.sf, Type %in% c("Compulsory", "Adaptive")),
               color = ~txPalLeg(Leg), weight = 1, opacity = 1, 
               label = ~htmlEscape(paste(Type, Transect)),
               popup = ~popup,
               group = "Planned Transects (Core)") %>% 
  addCircleMarkers(data = filter(wpts.sf, Type %in% c("Compulsory", "Adaptive")),
                   radius = 3, color = "#000414", stroke = FALSE, opacity = 1,
                   fillOpacity = 0.3, fillColor =  ~txPalLeg(Leg), 
                   label = ~htmlEscape(paste(Type, Waypoint)),
                   popup = ~popup,
                   group = "Planned Transects (Core)") %>%
  # Add ancillary planned transects
  addPolylines(data = filter(transects.sf, !Type %in% c("Compulsory", "Adaptive")),
               color = ~txPal(Type), weight = 1, opacity = 1,
               popup = ~popup,
               group = "Planned Transects (Ancillary)") %>%
  addCircleMarkers(data = filter(wpts.sf, !Type %in% c("Compulsory", "Adaptive")),
                   radius = 3, color = ~txPal(Type), stroke = FALSE, opacity = 1,
                   popup = ~popup,
                   group = "Planned Transects (Ancillary)") %>%
  # Add scale bar
  addScaleBar(position = "bottomright") %>%
  # Add map coordinates
  addMouseCoordinates() %>% 
  # Add measurement tool
  addMeasure(primaryLengthUnit = "miles", secondaryLengthUnit = "km",
             primaryAreaUnit = "sqmiles", secondaryAreaUnit = "sqmeters",
             position = "topleft") %>% 
  # Add layer controls
  addLayersControl(
    baseGroups = c("Esri.NatGeoWorldMap", "Esri.OceanBasemap",
                   "Esri.WorldImagery","Esri.WorldStreetMap","OpenStreetMap.Mapnik"),
    overlayGroups = c("Canada MPAs","Mexico MPAs","U.S. MPAs","State MPAs",
                      "State Waters", "Exclusive Economic Zones", 
                      "Bathymetry Contours",
                      "Planned Transects (Core)", "Planned Transects (Ancillary)",
                      "Vessel Track", "Vessel Positions"),
    options = layersControlOptions(collapsed = FALSE)) %>% 
  fitBounds(imap.bounds$range.lon[1], imap.bounds$range.lat[1],
            imap.bounds$range.lon[2], imap.bounds$range.lat[2])

if (exists("nav.paths.sf"))
  i.map <- i.map %>% 
  # Add nav data
  addPolylines(data = nav.paths.sf, 
               color = ~txPalLeg(Leg), weight = 3, opacity = 0.5,
               # color = "#000414", weight = 1, 
               label = ~leg, group = "Vessel Track") %>%
  addMarkers(data = nav.now, label = ~label, popup = ~popup, icon = ~oceanIcons["ship"],
             group = "Vessel Positions")

# Set visible layers
i.map <- i.map %>%  
  hideGroup(c("State Waters", "Exclusive Economic Zones",
              "Canada MPAs","Mexico MPAs","Planned Transects (Ancillary)"))
```

Survey progress for the **`r survey.name.long` (`r survey.name`)**

**Last updated:** `r format(Sys.time(), "%F %T", tz = "America/Los_Angeles", usetz = TRUE)`

**Disclaimer:** ***trackSurvey*** is a real-time tool for viewing survey progress relative to plans and objectives established at the beginning of the survey. Figures and tables in this report are constructed from data collected during the survey at various stages of QA/QC. **Therefore, this website may contain errors and is subject to change without notice.**

# Survey progress map

The map below shows the the projected (left) versus actual (right) survey progress. **Projections** are based on various factors including (but not limited to): planned days-at-sea (DAS), average vessel speed (e.g., `r survey.speed` kn), latitude-specific day length, total nautical miles of planned transects, and non-survey transit distances. **Actual** progress shows transects (or portions thereof) that have been sampled to-date. The semitransparent lines show the vessel's path for each leg of the survey. The white diamond shows the latest known location of the ship as of **`r with_tz(nav.now$time, tz = "America/Los_Angeles")` PDT** (`r nav.now$time` UTC). The outline of the diamond indicates the current leg of the survey.

```{r survey-progress-map}
#| fig-height: 10
#| fig-width: 10 

# Print plan map
map.plan.comp
```

# Survey progress plot
## By cumulative distance

Cumulative acoustic distance sampled (km) by *`r survey.vessel`* (pink line) relative to the cumulative planned acoustic transect distances (gray line). The vertical and horizontal dashed lines represent the expected date and cumulative distance at the end of each planned survey leg, respectively. Ideally, the actual progress line would align with the intersection of the two dashed lines at the end of a given survey leg.

```{r survey-progress-plot-distance}
#| fig-height: 7
#| fig-width: 12 

# Print plan plot
vessel.progress.plot.dist
```

## By latitude

Actual survey progress by *`r survey.vessel`* (pink line) relative to the original plan (gray line). Points filled [**black**]{style="color:black"} indicate dates when acoustic/trawl sampling occurred (i.e., survey days), points filled [**blue**]{style="color:blue"} indicate days in port (i.e., when speeds are less than 1 kn, for example when the ship is tied-up), and open points indicate transit days or days on the water with no acoustic transect data collected (i.e., no acoustic data collected). The vertical and horizontal dashed lines represent the expected date and latitude at the end of each planned survey leg, respectively. Ideally, the actual progress line would align with the intersection of the two dashed lines at the end of a given survey leg.

```{r survey-progress-plot}
#| fig-height: 7
#| fig-width: 12 

# Print plan plot
vessel.progress.plot
```

```{r summarize-das}
# Summarise planned DAS
das.planned <- data.frame(
  Leg = seq(1, length(leg.days)),
  planned.start = leg.ends[seq(1, length(leg.ends), 2)],
  planned.end = leg.ends[seq(2, length(leg.ends), 2)]) %>% 
  mutate(das.planned = as.numeric(planned.end - planned.start) + 1)

# Summarise actual DAS
das.actual <- nav %>% 
  # Calculate legs using the actual leg dates
  mutate(leg2 = cut(as.numeric(date(time)), 
                    as.numeric(leg.ends),
                    labels = FALSE)) %>% 
  filter(!is.na(Leg), leg2 %% 2 != 0) %>% 
  # filter(SOG > 1) %>% 
  group_by(leg2) %>% 
  summarise(
    actual.start = min(date(time)),
    actual.end = max(date(time)),
    das.actual = n_distinct(date(time))) %>% 
  mutate(Leg = seq_along(leg2)) %>% 
  select(Leg, everything(), -leg2)

# Load acoustic data, to summarise actual survey days
# load(here("Output/nasc_cps.Rdata"))

das.survey <- nasc.cps %>% 
  # Calculate legs using the actual leg dates
  mutate(leg2 = cut(as.numeric(date(datetime)), 
                    as.numeric(leg.ends),
                    labels = FALSE)) %>% 
  filter(leg2 %% 2 != 0) %>% 
  group_by(leg2) %>% 
  summarise(
    survey.start = min(date(datetime)),
    survey.end = max(date(datetime)),
    das.survey = n_distinct(date(datetime))) %>% 
  mutate(Leg = seq_along(leg2)) %>% 
  select(Leg, everything(), -leg2)

# Export data used to compute days doing surveys
nasc.survey <- nasc.cps %>% 
  # Calculate legs using the actual leg dates
  mutate(leg2 = cut(as.numeric(date(datetime)), 
                    as.numeric(leg.ends),
                    labels = FALSE)) %>% 
  filter(leg2 %% 2 != 0, !lat == 999, !long == 999) %>% 
  mutate(Leg = as.numeric(as.factor(leg2)),
         transect.num = as.numeric(
           substr(transect, 1,3))) %>% 
  select(Interval:time, datetime, transect, transect.num, Leg) 

das.summary <- das.planned %>% 
  left_join(das.actual) %>% 
  left_join(das.survey) %>% 
  mutate(delta.das = das.planned - das.actual,
         delta.survey = das.planned - das.survey)

# Save data frame with transect goals
tx.goals <- data.frame(
  Leg = seq_along(tx.breaks)-1,
  transect_goal = tx.breaks) %>% 
  slice(-1)

# Export for others
dir_create(here("Output/trackSurvey"))
write_csv(nav, 
          file = here("Output/trackSurvey/trackSurvey_nav.csv"))
write_csv(nasc.survey, 
          file = here("Output/trackSurvey/trackSurvey_nav_survey.csv"))
write_csv(das.planned, 
          file = here("Output/trackSurvey/trackSurvey_days_planned.csv"))
write_csv(das.actual, 
          file = here("Output/trackSurvey/trackSurvey_days_actual.csv"))
write_csv(das.survey, 
          file = here("Output/trackSurvey/trackSurvey_days_survey.csv"))
write_csv(das.summary, 
          file = here("Output/trackSurvey/trackSurvey_days_summary.csv"))
write_csv(tx.goals, 
          file = here("Output/trackSurvey/trackSurvey_transect_goals.csv"))

```

# Progress summaries so far

This table is meant to summarize planned versus realized survey effort (i.e., days at sea, or DAS). **Note:** these metrics are still under development and may not be completely accurate.

## Planned vs. actual effort

How many days has the ship been on the water (i.e., actual) versus how many days were planned? This is derived from the entire vessel track, regardless of whether or not we were conducting a survey (i.e., conducting acoustic sampling along planned transects), including transits and other vessel movements. [**Yellow**]{style="background-color:yellow"} cells may be a reason for concern. [**Red**]{style="background-color:red"} cells indicate potentially significant differences between the number of planned versus actual values and should be investigated further.

```{r effort-summary}
# Subset summary to only include legs that have been
das.summary.sub <- das.summary %>% 
  filter(!is.na(actual.start))

# Summarize the planned vs. actual survey DAS
das.summ.planned <- das.summary %>% 
  filter(!is.na(actual.start))

das.summary %>% 
  filter(!is.na(actual.start)) %>%
  filter(!is.na(survey.start)) %>%
  mutate(status = case_when(
    planned.end < Sys.Date() ~ "Completed",
    TRUE ~ "Ongoing"
  )) %>% 
  mutate(
    actual.start = cell_spec(
      actual.start, knitr.format, color = "black",
      background = ifelse(actual.start >= planned.start, "red", "white")),
    actual.end = cell_spec(
      actual.end, knitr.format, color = "black",
      background = case_when(
        status == "Ongoing" ~ "yellow",
        actual.end < planned.end ~ "red",
        TRUE ~ "white")),
    survey.start = cell_spec(
      survey.start, knitr.format, color = "black",
      background = ifelse(survey.start > planned.start + days(das.survey.transit),
                          "red", "white")),
    survey.end = cell_spec(
      survey.end, knitr.format, color = "black",
      background = ifelse(survey.end < planned.end - days(das.survey.transit),
                          "red", "white")),
    status = cell_spec(
      status, knitr.format, bold = TRUE,
      background = ifelse(status == "Completed", "white", "green"),
      color      = ifelse(status == "Completed", "black", "white")),
    das.actual = cell_spec(
      das.actual, knitr.format, color = "black",
      background = case_when(
        das.actual == das.planned ~ "green", 
        das.actual < das.planned & das.actual > das.planned - 2 ~ "yellow",
        TRUE ~ "red")),
    das.survey = cell_spec(
      das.survey, knitr.format, color = "black",
      background = case_when(
        das.survey >= das.planned - 2*das.survey.transit ~ "green",
        das.survey < das.planned - 2*das.survey.transit & 
          das.survey > das.planned - 4*das.survey.transit ~ "yellow",
        TRUE ~ "red"),
      bold = if_else(Sys.Date() < planned.end, TRUE, FALSE))) %>%
  select(Leg, planned.start, actual.start, planned.end, actual.end, 
         das.planned, das.actual, das.survey, status) %>%
  kable(format = knitr.format,booktabs = TRUE, escape = FALSE,
        col.names = c("Leg", 
                      "Planned", "Actual", 
                      "Planned", "Actual", 
                      "Planned", "Actual", "Survey",
                      "Status")) %>% 
  add_header_above(c(" " = 1, "Start date" = 2, "End date" = 2, 
                     "Days at Sea" = 3, " " = 1)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), 
                full_width = FALSE) 

```

According to our calculations, a total of **`r sum(das.summary.sub$das.actual, na.rm = T)`** DAS were accomplished of the `r sum(das.summary.sub$das.planned, na.rm = T)` DAS scheduled, resulting in **`r sum(das.summary.sub$delta.das, na.rm = T)`** fewer DAS than planned.

**NOTE:** If the Status is "[**Ongoing**]{style="background-color:green"}" then these calculations will not accurately reflect the actual results.

```{r das-summary,eval=FALSE}
# Subset summary to only include legs that have been
das.summary.sub <- das.summary %>% 
  filter(!is.na(actual.start))

das.summary %>% 
  filter(!is.na(actual.start)) %>% 
  mutate(
    actual.start = cell_spec(
      actual.start, knitr.format, color = "black",
      background = ifelse(actual.start >= planned.start, "red", "white")),
    actual.end = cell_spec(
      actual.end, knitr.format, color = "black",
      background = ifelse(actual.end < planned.end, "red", "white")),
    delta.das = cell_spec(
      delta.das, knitr.format, color = "black",
      background = ifelse(delta.das < das.planned, "red", "white"))) %>%
  select(Leg, planned.start, actual.start, planned.end, actual.end, 
         das.planned, das.actual, delta.das) %>%
  kable(format = knitr.format,booktabs = TRUE, escape = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"), 
                full_width = FALSE) 
```

```{r survey-summary, eval=FALSE}
# ## Planned vs. actual **survey** days at sea (DAS)
# 
# How many days has the ship been conducting acoustic transect sampling versus how many days were planned? This is derived from the acoustic data collected along planned transects, and excludes transits between transects and to/from ports of call. Red cells indicate differences between scheduled and accomplished values. These values allow for **`r das.survey.transit`** DAS for transit at the beginning and end of each survey leg, when acoustic transect sampling may not actually be conducted. Values in [**red**]{style="color:red"} below are a potential reason for concern.

# Summarize the planned vs. actual survey DAS
das.summ.planned <- das.summary %>% 
  filter(!is.na(actual.start))

das.summary %>% 
  filter(!is.na(actual.start)) %>%
  filter(!is.na(survey.start)) %>%
  mutate(
    survey.start = cell_spec(
      survey.start, knitr.format, color = "black",
      background = ifelse(survey.start > planned.start + days(das.survey.transit),
                          "red", "white")),
    survey.end = cell_spec(
      survey.end, knitr.format, color = "black",
      background = ifelse(survey.end < planned.end - days(das.survey.transit),
                          "red", "white")),
    delta.survey = cell_spec(
      delta.survey, knitr.format, color = "black",
      # 2 DAS seems like  reasonable number of transit days at the beginning/end
      # of a survey leg, so anything greater than that is reason for concern.
      background = ifelse(delta.survey > das.survey.transit*2, "red", "white"))) %>%
  select(Leg, planned.start, survey.start, planned.end, survey.end, 
         das.planned, das.survey, delta.survey) %>%
  kable(format = knitr.format,booktabs = TRUE, escape = FALSE) %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"), 
                full_width = FALSE) 

# According to our calculations, a total of **`r sum(das.summary.sub$das.survey, na.rm = T)`** DAS were accomplished of the `r sum(das.summary.sub$das.planned, na.rm = T)` DAS scheduled, resulting in **`r sum(das.summary.sub$delta.survey, na.rm = T)`** fewer days of acoustic sampling than what was scheduled.
```

```{r survey-summary-rev,eval=FALSE}
# ## Planned vs. actual **survey** days at sea (DAS)
# 
# How many days has the ship been conducting acoustic transect sampling versus how many days were planned? This is derived from the acoustic data collected along planned transects, and excludes transits between transects and to/from ports of call. Red cells indicate differences between scheduled and accomplished values. These values allow for **`r das.survey.transit`** DAS for transit at the beginning and end of each survey leg, when acoustic transect sampling may not actually be conducted. Values in [**red**]{style="color:red"} below are a potential reason for concern.

# Summarize the planned vs. actual survey DAS
das.summ.planned <- das.summary %>% 
  filter(!is.na(actual.start))

das.summary %>% 
  filter(!is.na(actual.start)) %>%
  filter(!is.na(survey.start)) %>%
  mutate(
    actual.start = cell_spec(
      actual.start, knitr.format, color = "black",
      background = ifelse(actual.start >= planned.start, "red", "white")),
    actual.end = cell_spec(
      actual.end, knitr.format, color = "black",
      background = ifelse(actual.end < planned.end, "red", "white")),
    survey.start = cell_spec(
      survey.start, knitr.format, color = "black",
      background = ifelse(survey.start > planned.start + days(das.survey.transit),
                          "red", "white")),
    survey.end = cell_spec(
      survey.end, knitr.format, color = "black",
      background = ifelse(survey.end < planned.end - days(das.survey.transit),
                          "red", "white")),
    das.actual = cell_spec(
      das.actual, knitr.format, color = "black",
      background = ifelse(das.actual < das.planned, "red", "white")),
    das.survey = cell_spec(
      das.survey, knitr.format, color = "black",
      background = ifelse(das.survey < das.planned - 2*das.survey.transit, "red", "white"))) %>%
  select(Leg, planned.start, actual.start, planned.end, actual.end, 
         das.planned, das.actual, das.survey) %>%
  kable(format = knitr.format,booktabs = TRUE, escape = FALSE,
        col.names = c("Leg", 
                      "Planned", "Actual", 
                      "Planned", "Actual", 
                      "Planned", "Actual", "Survey")) %>% 
  add_header_above(c(" " = 1, "Start date" = 2, "End date" = 2, "Days at Sea" = 3)) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), 
                full_width = FALSE) 

# **NOTE:** If the `planned.end` date for the current leg is greater than **`r date(Sys.time())`**, then the survey leg is still ongoing and these calculations will not accurately reflect the actual results.
```

# Interactive data explorer

An interactive map showing the path of the vessel (heavy lines) relative to planned acoustic transects (lighter lines). The colors of the lines indicate the expected leg that a planned transect would be sampled and the actual leg the vessel completed sampling. Hover-over or click on some map objects (e.g., the ship icon) for more info.

```{r interactive-map, out.height="12in", out.width="100%"}
# Display map
i.map
```

For questions about this page, please contact Kevin Stierhoff ([kevin.stierhoff\@noaa.gov](mailto:kevin.stierhoff@noaa.gov)).
